---
title: Tutorial 6  
author: STAT 3013/8027
output: pdf_document
---

---
\large

1. Consider the following simple linear regression model:

\begin{eqnarray*}
Y_i &=& \beta_0 + \beta_1 x_i + \epsilon_i \\
\epsilon_i &\sim iid& \ n(0, \sigma^2), \ \ i=1, \ldots, n. 
\end{eqnarray*}

**Ans.** See the handwritten pages for $\hat{\beta_0}$ and $\hat{\beta_1}$, as well as their expectations and variances.

```{r}
gdp <- read.csv("gdp2013.csv", header=T)
labor <- read.csv("labor2013.csv", header=T)
D <- merge(gdp, labor, by=c("Country.Name", "Country.Code"))
dim(D)
D <- na.omit(D)
dim(D)
names(D)[3:4] <- c("gdp", "labor")

##
y <- log(D$gdp)
x <- log(D$labor)


##
S.xy <- sum ( (y-mean(y))* (x-mean(x)) )
S.xx <- sum ( (x-mean(x))^2 )
beta.1.hat <- S.xy/S.xx
beta.1.hat

beta.0.hat <- mean(y) - beta.1.hat*mean(x)
beta.0.hat

mod <- lm(y ~ x)
summary(mod)
```



2. Least-squares estimates for a categorical regression model.  **Ans.** See the handwritten pages for the derivation of $\hat{\mu}_j$.  We found:

    $$\hat{\mu}_j = \frac{1}{n} \sum_{i=1}^n y_{i,j} = \bar{y}_{.,j}$$
    
```{r}
y <- c(21, 23, 19, 24, 25, 23, 19, 21, 20, 18, 
       22, 20, 15, 10, 13, 14, 11, 15)
x <- as.factor(c(rep("0.5",6),  rep("1.0",6), rep("1.5",6)))   
plot(y ~ x)
```


Note:  In our model we assume every $y_{ij}$ has the same variability.  A typical concern is whether the variability within each group is similar across the groups.  As a quick check we can examine the box part of the box plots to see if they are reasonably similar in spread.  It seems roughly OK here, although group 2 has a slightly smaller spread comparatively.

```{r}
mu.hat <- tapply (y, x ,mean)    
mu.hat
````    
