---
title: "Limiting spectral distributions"
output:
  html_document:
---

First, we load some packages.
```{r}
library(Matrix)   # matrix operations
library(MASS)     # Multivariate Normal Distribution
```


Sample a covariance based on $\mathbb{S}_n = \frac{1}{n} \mathbb{X} \mathbb{X^*}$ where $\mathbb{X}$ is a $p \times n$ matrix with Gaussian entries with mean zero and variance 1. This gives a matrix $\mathbb{S}_n$ of size $p\times p$.

```{r}
p <- 50
n <- 500
X <- matrix(rnorm(p*n), p, n)
Sn <- X %*% t(X) / n
dim(Sn)
```

Calculate the ratio $y=p/n$
```{r}
p/n
```

Calculate the eigenvalues.
```{r}
e<-eigen(Sn)
L<-e$values
```

We have $p$ eigenvalues.

```{r}
length(L)
```

Print out the eigenvalues, notice that are all real numbers.
```{r}
L
```

Draw a horizontal line at every eigenvalue.
```{r}
plot(c(0,1.5*max(L)), c(0,1), type='n', xlab='x', ylab='')
abline(v=L, col=2)
```

```{r}
F_ <- function(x) {
  total <- 0.
  for (i in 1:p) {
    if (L[i] <= x) {
      total <- total + 1
    }
  }
  return (total/p)
}
F <- Vectorize(F_)
```

```{r}
plot(F, from=0., to=2*max(L), main="Empirical CDF")
```

Plot a histogram of the eigenvalues.
```{r}
hist(L, breaks=p/3, xlim=c(0,1.2*max(L)), freq=FALSE, col=1, main='')
```