---
title: "Tutorial 2 Solutions"
author: "STAT 3013/8027"
output: pdf_document
---
---
\large

1.  Rice: 5.1, 5.13, 5.16  **Ans.** See the handwritten pages.

* **5.13 R code:**


```{r}
##
set.seed(10)
x <- rbinom(60, 1, 0.5)
x[x==0] <- -1
x <- x*50

w <- cumsum(x)

##
plot(w, type="l", lwd=3, main="A Realization of the Drunkard's Walk")


## Let's examine the sampling distirbution of W
set.seed(10)
S <- 10000
W <- rep(0, S)

for(s in 1:S){
  x <- rbinom(60, 1, 0.5)
  x[x==0] <- -1
  x <- x*50
  W[s] <- sum(x)
}

##
mean(W)
var(W)

##
hist(W, col="yellow")
```

* **5.19 (a)**

Let's first work out the exact answer.

\begin{eqnarray*}
\int_0^1 cos(2 \pi x) dx &=& \frac{1}{2 \pi} sin(2 \pi x) \big\vert_0^1 \\
&=& 0
\end{eqnarray*}

  + Note.  Suppose we consider $U \sim \textrm{Uniform(0,1)}$.  Let's look at the $E[U]$.
  
  $$E[U] = \int_0^1 u f(u) du = \int_0^1 u \ du $$
  
  + Now let's look at the expected value of the function: $cos(2 \pi u)$:
  
  $$E[g(U)] = \int_0^1 cos(2 \pi u) \ du $$

**So our integral of interest is the expected value of the function.  We can approximate that via simulation.**

$$\hat{I}(g(u)) = \frac{1}{S}\sum_{1}^S cos(2 \pi u)$$

```{r}
set.seed(1001)
S1 <- 100

u <- runif(S1)
g.u <- cos(2*pi*u)
mean(g.u)
```

```{r}
set.seed(1001)
S2 <- 1000

u <- runif(S2)
g.u <- cos(2*pi*u)
mean(g.u)
```


+ Now let's plot for increasing values of $S$:

```{r}
S <- 50:5000
g.u <- rep(0, length(S))

c <- 1
for(s in S){
g.u[c] <- mean(cos(2*pi* runif(s)))
c <- c+1
}

plot(S, g.u, type="l", lwd=3, col="blue") 
abline(h=0, col="red", lwd=4)
```

+ As we can work out the $E[g(u)]$ and the $V[g(u)]$ we could also use the CLT theorem for calculations based on the $\frac{1}{S}\sum_{1}^S cos(2 \pi u)$.


* **5.19 (b) R code**

  + For this question we can't work out an analytical solution (some type of approximation must be performed).

\begin{eqnarray*}
E[I(g(u))] &=& \int_0^1 cos(2 \pi u^2) du\\
&\approx& \frac{1}{S} \sum_1^S  cos(2 \pi u^2)
\end{eqnarray*}

```{r}
set.seed(1001)
S1 <- 100

u <- runif(S1)
g.u <- cos(2*pi*u^2)
mean(g.u)
```

```{r}
set.seed(1001)
S2 <- 1000

u <- runif(S2)
g.u <- cos(2*pi*u^2)
mean(g.u)
```

```{r}
S <- 50:5000
g.u <- rep(0, length(S))

c <- 1
for(s in S){
g.u[c] <- mean(cos(2*pi* runif(s)^2))
c <- c+1
}

plot(S, g.u, type="l", lwd=3, col="blue") 
```  

