---
title: "tutnotes3"
author: "Rui Qiu"
date: '2017-08-09'
output: html_document
---

## two-way ANOVA model
two commonly used constraints:

- the 'baseline' or 'control group structure' $\tau_1=\alpha_1=0$ or
- the 'grand mean' constraint $\sum^I_{i=1}\tau_i = \sum^J_{j=1}\alpha_j=0$

- also can use matrix form to represent (with the idea of multiple regression)

a two-way ANOVA model

\[Y_{ijk}=\mu+\tau_i+\alpha_j+\epsilon_{ijk}\]

```{r, eval=F}
aov(Y~X)
aov(Y~X1+X2)
```

## Q2
load data

```{r}
resp <- c(9, 11, 20, 19, 23, 19, 12, 10, 17, 18, 21, 20, 16, 24, 28, 25, 15, 25)
btyp <- c(rep("O-", 2), rep("O+", 4), rep("A-", 2), rep("A+", 4), "B-", rep("B+", 3), "AB-", "AB+")
```


###(a)
Bfactor: B-, B+, AB-, AB+
Rhfactor: O+, A+, B+, AB+

set up the level factors:
```{r}
Bfact <- iflelse(btype=="B-", "B", "notB")
Bfact <- iflelse(btype=="B+", "B", Bfact)
Bfact <- iflelse(btype=="AB-", "B", Bfact)
Bfact <- iflelse(btype=="AB+", "B", Bfact)
Rhfact <- iflelse(btype=="O+", "+", "-")
Rhfact <- iflelse(btype=="A+", "+", Rhfact)
Rhfact <- iflelse(btype=="B+", "+", Rhfact)
Rhfact <- iflelse(btype=="AB+", "+", Rhfact)
```


Fit 2-way ANOVA models:
```{r}
btyp.aov <- aov(resp ~ Bfact + Rhfact)
summary(btyp.aov)
```

Alternatively, use `lm()`
```{r}
btyp.aov <- lm(resp ~ Bfact + Rhfact)
summary(btyp.aov)
```

> Note: order changed, the results won't change a lot (for this case). It's kind of special.

### (b)
```{r}
```

### (c)


