---
title: "Tutorial 1 Solutions"
author: "STAT 3013/8027"
output: pdf_document
---
---
\large

1.  Rice: 2.31, 3.18.  **Ans.** See the handwritten pages.

2.  Consider the following:

      a. Visually display the data and discuss.  Try taking the natural log of the data (when statisticians say "log" they mean natural log).
      b. Compute a six number summary of the data.
      c. Based on the ``box plot rule'', determine if there are any outliers.  Which countries are outliers?  To use the rule examine the following: Are any values in the data below the $1^{\textrm{st}}$ Quartile - 1.5 IQR?  Are any values in the data above the $3^{\textrm{rd}}$ Quartile + 1.5 IQR?  IQR is the inter-quartile range.
      d. Let $Y$ = log(GDP).  Suppose $Y \sim \textrm{normal}(\mu, \sigma^2)$.  What is your best guess for $\mu$ and $\sigma^2$ as functions of $Y$ (call these $T_1$ and $T_2$)?  What are the means (expected values) of $T_1$ and $T_2$? 
    
**Ans.** First let's load in the data.  Note that I removed the missing values (NA's).  Missing data is an extensive and important topic in statistics; as such be careful about removing missing data.  At the very least, discuss your full sample of data and then which cases were removed due to missing data.
```{r}
## GDP
gdp2013 <- read.table("gdp2013.txt", header=TRUE)
D <- gdp2013
D <- na.omit(D)
gdp <- D$Y2013
```


The histogram is unimodal, right skewed and appears to have some outliers.
```{r}
## visual display
hist(gdp, col="yellow")
```

If we take the log of the data it is far more symmetric.
```{r}
## visual display
hist(log(gdp), col="yellow")
```

Let's get a six number summary (using the logged data)
```{r}
log.gdp <- log(gdp)
summary(log.gdp)
```


Let's get the quartiles and the IQR.
```{r}
##
Q <- quantile(log.gdp, prob=c(0.25, 0.5, 0.75), type=6) 
Q

IQR <- unname(Q[3]-Q[1]) 
IQR

```


Let's identify outliers based on the box plot method.
```{r}
##
high.outliers <- log.gdp[log.gdp > Q[3] + 1.5*IQR]
low.outliers <- log.gdp[log.gdp < Q[1] - 1.5*IQR]

high.outliers
low.outliers

# let's find the countries that have high outliers
D[log.gdp >= min(high.outliers),1]

# let's find the countries that have high outliers
D[log.gdp <= max(low.outliers),1]


# let's make a boxplot.
boxplot(log.gdp, col="yellow")
```
    
* If we assume the data $Y_1, \ldots, Y_n \sim iid \ \ n(\mu, \sigma^2)$ then reasonable guesses for the population mean and variance are the sample mean and variances.  So we have:

$$\bar{y} = `r mean(log.gdp)` = \hat{\mu}$$

$$S^2 = `r var(log.gdp)` = \hat{\sigma}^2$$

A nice property of these estimators is that $E[ \hat{\mu} ]  = E[\bar{Y}] = \mu$ and $E[ \hat{\sigma}^2 ]  = E[S^2] = \sigma^2$.  We proved the latter in lecture and for fun let's do the former:

\begin{eqnarray*}
E[ \hat{\mu} ]  = E[\bar{Y}] &=& E[ (1/n) (Y_1 + \cdots + Y_n) ] \\
&=& (1/n) E[(Y_1 + \cdots + Y_n) ] \\
&=& (1/n) ( E[Y_1] + \cdots + E[Y_n] )  \\
&=& (1/n) n E[Y_1] = (1/n) n \mu = \mu
\end{eqnarray*}




    
