---
title: "tutnotes2"
author: "Rui Qiu"
date: '2017-08-02'
header-includes:
   - \usepackage{amssymb}
   - \usepackage{amsmath}
output: html_document
---

## Recap

The general form of linear model is

\[Y=\beta_0+\beta_1X_1+\beta_2X_2+\beta_3X_3+\epsilon\]

In matrix/vector representation...

Use **least square estimation** we have:

\[\hat{\beta}=(\mathbf X^T\mathbf X)^{-1}\mathbf X^T\mathbf y\]

## Inference

We have $\epsilon \sim N(0,\sigma^2\mathbf{I})$

Since $\mathbf{y}=\mathbf{X}\beta+\epsilon$, then $\mathbf{y}\sim N(\mathbf{X}\beta, \sigma^2\mathbf{I})$.

CI.

## Hypothesis test

## one-way ANOVA

***
\[
X=\begin{cases} 
      \text{level 1} & \mu_1 \\
      \text{level 2} & \mu_2 \\
      \text{level 3} & \mu_3
   \end{cases}
\]

$H_0: \mu_1=\mu_2$

$H_0: \mu_3-\mu_2 = \mu_2-\mu_1$

$H_0: h_1\mu_1+h_2\mu_2+h_3\mu_3=0$

\[Var(\sum^k_{i=1}h_i\bar{Y}_i)=\sum^k_{i=1}h_i^2Var(\bar{Y}_i)=\sigma^2\sum^k_{i=1}\frac{h_i^2}{n_i}\]

***
## problem 1
### (a)
```{r}
prod <- read.csv("productivity.csv", header=T, sep="\ ")
attach(prod)
plot(RandD, prodscre)
# not a problem (heteroscedasticity)
```

### (b)
```{r}
m1 <- lm(prodscre ~ RandD)
anova(m1)
par(mfrow=c(2,2))
plot(m1, which=c(1,2,4,5))
# another way to do so (one-way anova):
aov(m1) # the output might seem in different forms, but the results are genuinely the same
```

### (c)
test $\mu_{\text{med}} - \mu_{\text{low}} = \mu_{\text{high}} - \mu_{\text{med}}$

i.e. $\mu_{\text{high}} - 2\mu_{\text{med}} + \mu_{\text{low}} = 0$

\[
\begin{cases} 
      \mu_{\text{high}} &= \beta_0 \\
      \mu_{\text{low}} &= \beta_0 + \beta_1 \\
      \mu_{\text{med}} &= \beta_0 + \beta_2 \\
   \end{cases}
\]

$H_0: \beta_1=2\beta_2$, i.e. $-2\beta_2+\beta_1=0$

$\vec{h}=(0,1,-2)$

Then apply the formula above (in the very beginning), with `summary(m1)$sigma` we get the $\sigma$ value so that we have the variance.

***
## Q2
### (c)
the $\mu$ here actually does not mathematically mean the average, but the influence of two levels of factors, i.e. the null hypothesis $H_0$ means the influence of B factor is the same as the influence of Rh factor.

\[H_0:\frac{\mu_{B+}+\mu_{O+}}{2}-\frac{\mu_{B-}+\mu_{O-}}{2}=\frac{\mu_{B+}+\mu_{B-}}{2}-\frac{\mu_{O+}+\mu_{O-}}{2}\]
